---
title: "Assignment1"
output: html_document
date: "2023-05-09"
---

Import libraries

```{r}
library(tidyverse)
library(reshape2)
library(purrr)
library(dplyr)
library(ggplot2)
```

**Create agents**


1. Random agent
2. Alternating agent
3. Win-shift-lose-stay agent

**1. Random agent**

```{r}
RandomAgent_f <- function(rate){
  choice <- rbinom(1, 1, rate)
  return(choice)
}

```


**2. Alternating agent**

```{r}
#Defining agent function
AlternatingAgent_f <- function(prevChoice) {
  if (prevChoice == 1) {      #If prevChoice = 1 -> 1 - prevChoice
    choice = 1 - prevChoice
  }
  else if (prevChoice == 0) { #If prevChoice = 0 -> 1 + prevChoice
    choice = 1 + prevChoice
  }
  return(choice)
}
```



**3. Win-shift-lose-stay agent with bias**

```{r}
WSLSAgentBias_f <- function(prevChoice, Feedback, bias) {
  if (Feedback == 0) {      # If feedback = 0 (loss), stay
    choice = prevChoice
  }
  else if (Feedback == 1) { # If feedback = 1 (win), agent has a certain bias to shift, stay otherwise
    if (runif(1) < bias) {
      choice = 1 - prevChoice  # Shift
    } else {
      choice = prevChoice  # Stay
    }
  }
  return(choice)
}
```



**1. Alternating agent vs Random agent**

```{r}

rate = 0.5

# Define number of trials
trials = 120

# Define empty vectors for agents
Self <- rep(NA, trials)   # Alternating agent
Other <- rep(NA, trials)  # Random agent

# Define random first choice for Alternating agent
Self[1] <- RandomAgent_f(rate)

# Run simulation
for (t in seq(trials)){Other[t] <- RandomAgent_f(rate)}
for (i in 2:trials){
    Self[i] <- AlternatingAgent_f(Self[i-1])
}

# Visualizing 
# The choices
df <- tibble(Self, Other, trial = seq(trials), Feedback = as.numeric(Self==Other))
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, Self)) + geom_line(color="blue", aes(trial, Other))

# The performance
df <- df %>% mutate(
cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback), cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, cumulativerateSelf)) + geom_line(color="blue", aes(trial, cumulativerateOther))

```

**2. Win-Shift-Lose-Stay with bias vs Random agent**

```{r}
# Define rate
rate <- 0.5

# Define the probability of shifting after a win
bias <- 0.7

# Define number of trials
trials = 120

# Define empty vectors for agents
Self <- rep(NA, trials)   #Win-shift-lose-stay agent
Other <- rep(NA, trials)  #Random agent

# Define random first choice for agents
Self[1] <- RandomAgent_f(rate)
Other[1] <- RandomAgent_f(rate)

# Run simulation
for (i in 2:trials) {
  if (Self[i-1] == Other[i-1]) {
    Feedback = 1   # Win
  } else {
    Feedback = 0   # Loss
  }
  Self[i] <- WSLSAgentBias_f(Self[i-1], Feedback, bias)
  Other[i] <- RandomAgent_f(rate)
}

#Visualizing 
#The choices
df <- tibble(Self, Other, trial = seq(trials), Feedback = as.numeric(Self==Other))
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, Self)) + geom_line(color="blue", aes(trial, Other))

#The performance
df <- df %>% mutate(
cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback), cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, cumulativerateSelf)) + geom_line(color="blue", aes(trial, cumulativerateOther))

```

**3. Win-Shift-Lose-Stay with bias vs Alternating agent**

```{r}
# Define rate
rate <- 0.5

# Define the probability of shifting after a win
bias <- 0.7 

# Define number of trials
trials = 120

#Define empty vectors for agents
Self <- rep(NA, trials)
Other <- rep(NA, trials)

#Define random first choice for agents
Self[1] <- RandomAgent_f(rate)
Other[1] <- RandomAgent_f(rate)

#Run simulation
for (i in 2:trials){
  if (Self[i-1] == Other[i-1]){
    Feedback = 1   #Win
  } else {
    Feedback = 0   #Loss
  }
  Self[i] <- WSLSAgentBias_f(Self[i-1], Feedback, bias)
  Other[i] <- AlternatingAgent_f(Other[i-1])
}


#Visualizing 
#The choices
df <- tibble(Self, Other, trial = seq(trials), Feedback = as.numeric(Self==Other))
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, Self)) + geom_line(color="blue", aes(trial, Other))

#The performance
df <- df %>% mutate(
cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback), cumulativerateOther = cumsum(1-Feedback) / seq_along(Feedback) )
ggplot(df) + theme_classic() + geom_line(color="red", aes(trial, cumulativerateSelf)) + geom_line(color="blue", aes(trial, cumulativerateOther))
```




**1. Alternating agent vs random agent for 100 dyads**

```{r}
# Define number of trials and dyads
trials <- 120
dyads <- 100
rate <- 0.5

# Initialize a list to store the results of each dyad
results <- vector("list", dyads)

# Run simulation for each dyad
for (d in 1:dyads) {
  # Define empty vectors for agents
  Self <- rep(NA, trials)   # Alternating agent
  Other <- rep(NA, trials)  # Random agent

  # Define random first choice for agent
  Self[1] <- RandomAgent_f(rate)

  # Run simulation
  for (t in seq(trials)) {Other[t] <- RandomAgent_f(rate)}
  for (i in 2:trials) {
    if (Self[i-1] == Other[i-1]) {
      Self[i] <- AlternatingAgent_f(1)
    } else {
      Self[i] <- AlternatingAgent_f(0)
    }
  }
  
  # Store the results for this dyad in the list
  results[[d]] <- list(Self = Self, Other = Other)
}

# Convert results to a data frame
df <- purrr::map2_df(results, seq_along(results), ~{
  tibble(
    trial = seq_along(.x$Self),
    Self = .x$Self,
    Other = .x$Other,
    Feedback = as.numeric(.x$Self == .x$Other)
  ) %>% mutate(
    cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback),
    cumulativerateOther = cumsum(1 - Feedback) / seq_along(Feedback),
    dyad = paste("Dyad", .y)  # add a column for the dyad number
  )
})

# Compute the mean cumulative success rates
mean_df <- df %>%
  group_by(trial) %>%
  summarise(
    mean_cumulativerateSelf = mean(cumulativerateSelf),
    mean_cumulativerateOther = mean(cumulativerateOther)
  )

# Filter to the final trial for each dyad
final_trial_df <- df %>%
  group_by(dyad) %>%
  filter(trial == max(trial))

# Compute the mean of the final cumulative success rates
mean_final_cumulativerateSelf <- mean(final_trial_df$cumulativerateSelf)
mean_final_cumulativerateOther <- mean(final_trial_df$cumulativerateOther)

# Print the results
print(paste("Mean final cumulative success rate for Self: ", mean_final_cumulativerateSelf))
print(paste("Mean final cumulative success rate for Other: ", mean_final_cumulativerateOther))


# Plot the cumulative rates for each dyad in a single plot
ggplot(df, aes(trial)) + 
  geom_line(aes(y = cumulativerateSelf, color = dyad)) + 
  geom_line(aes(y = cumulativerateOther, color = dyad)) +
  geom_line(data = mean_df, aes(y = mean_cumulativerateSelf), color = "red", size = 1.5) +
  geom_line(data = mean_df, aes(y = mean_cumulativerateOther), color = "blue", size = 1.5) +
  theme_classic() +
  ggtitle("Alternating agent vs random agent") +
  theme(legend.position = "none")  # remove the legend for clarity

```


**2. Win-Shift-Lose-Stay agent with bias vs Random agent for 100 dyads**


```{r}
# Define number of trials and dyads
trials <- 120
dyads <- 100
rate <- 0.5
bias <- 0.7

# Initialize a list to store the results of each dyad
results <- vector("list", dyads)

# Run simulation for each dyad
for (d in 1:dyads) {
  # Define empty vectors for agents
  Self <- rep(NA, trials)   # Win-shift-lose-stay agent
  Other <- rep(NA, trials)  # Random agent

  # Define random first choice for agent
  Self[1] <- RandomAgent_f(rate)

  # Run simulation
  for (t in seq(trials)){Other[t] <- RandomAgent_f(rate)}
  for (i in 2:trials) {
  if (Self[i-1] == Other[i-1]) {
    Feedback = 1   # Win
  } else {
    Feedback = 0   # Loss
  }
  Self[i] <- WSLSAgentBias_f(Self[i-1], Feedback, bias)
  Other[i] <- RandomAgent_f(rate)
}
  
  # Store the results for this dyad in the list
  results[[d]] <- list(Self = Self, Other = Other)
}

# Convert results to a data frame
df <- purrr::map2_df(results, seq_along(results), ~{
  tibble(
    trial = seq_along(.x$Self),
    Self = .x$Self,
    Other = .x$Other,
    Feedback = as.numeric(.x$Self == .x$Other)
  ) %>% mutate(
    cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback),
    cumulativerateOther = cumsum(1 - Feedback) / seq_along(Feedback),
    dyad = paste("Dyad", .y)  # add a column for the dyad number
  )
})

# Compute the mean cumulative success rates
mean_df <- df %>%
  group_by(trial) %>%
  summarise(
    mean_cumulativerateSelf = mean(cumulativerateSelf),
    mean_cumulativerateOther = mean(cumulativerateOther)
  )

# Filter to the final trial for each dyad
final_trial_df <- df %>%
  group_by(dyad) %>%
  filter(trial == max(trial))

# Compute the mean of the final cumulative success rates
mean_final_cumulativerateSelf <- mean(final_trial_df$cumulativerateSelf)
mean_final_cumulativerateOther <- mean(final_trial_df$cumulativerateOther)

# Print the results
print(paste("Mean final cumulative success rate for Self: ", mean_final_cumulativerateSelf))
print(paste("Mean final cumulative success rate for Other: ", mean_final_cumulativerateOther))


# Plot the cumulative rates for each dyad in a single plot
ggplot(df, aes(trial)) + 
  geom_line(aes(y = cumulativerateSelf, color = dyad)) + 
  geom_line(aes(y = cumulativerateOther, color = dyad)) +
  geom_line(data = mean_df, aes(y = mean_cumulativerateSelf), color = "red", size = 1.5) +
  geom_line(data = mean_df, aes(y = mean_cumulativerateOther), color = "blue", size = 1.5) +
  theme_classic() +
  ggtitle("Win-shift-lose-stay agent vs random agent") +
  theme(legend.position = "none")  # remove the legend for clarity


```


**3. Win-Shift-Lose-Stay agent with bias vs Alternating agent for 100 dyads**


```{r}
# Define number of trials and dyads
trials <- 120
dyads <- 100
rate <- 0.5
bias <- 0.7

# Initialize a list to store the results of each dyad
results <- vector("list", dyads)

# Run simulation for each dyad
for (d in 1:dyads) {
  # Define empty vectors for agents
  Self <- rep(NA, trials)   # Win-shift-lose-stay agent
  Other <- rep(NA, trials)  # Alternating agent

  #Define random first choice for agents
  Self[1] <- RandomAgent_f(rate)
  Other[1] <- RandomAgent_f(rate)

  #Run simulation
  for (i in 2:trials){
    if (Self[i-1] == Other[i-1]){
      Feedback = 1   #Win
    } else {
      Feedback = 0   #Loss
    }
    Self[i] <- WSLSAgentBias_f(Self[i-1], Feedback, bias)
    Other[i] <- AlternatingAgent_f(Other[i-1])
  }
  
  # Store the results for this dyad in the list
  results[[d]] <- list(Self = Self, Other = Other)
}

# Convert results to a data frame
df <- purrr::map2_df(results, seq_along(results), ~{
  tibble(
    trial = seq_along(.x$Self),
    Self = .x$Self,
    Other = .x$Other,
    Feedback = as.numeric(.x$Self == .x$Other)
  ) %>% mutate(
    cumulativerateSelf = cumsum(Feedback) / seq_along(Feedback),
    cumulativerateOther = cumsum(1 - Feedback) / seq_along(Feedback),
    dyad = paste("Dyad", .y)  # add a column for the dyad number
  )
})

# Compute the mean cumulative success rates
mean_df <- df %>%
  group_by(trial) %>%
  summarise(
    mean_cumulativerateSelf = mean(cumulativerateSelf),
    mean_cumulativerateOther = mean(cumulativerateOther)
  )

# Filter to the final trial for each dyad
final_trial_df <- df %>%
  group_by(dyad) %>%
  filter(trial == max(trial))

# Compute the mean of the final cumulative success rates
mean_final_cumulativerateSelf <- mean(final_trial_df$cumulativerateSelf)
mean_final_cumulativerateOther <- mean(final_trial_df$cumulativerateOther)

# Print the results
print(paste("Mean final cumulative success rate for Self: ", mean_final_cumulativerateSelf))
print(paste("Mean final cumulative success rate for Other: ", mean_final_cumulativerateOther))


# Plot the cumulative rates for each dyad in a single plot
ggplot(df, aes(trial)) + 
  geom_line(aes(y = cumulativerateSelf, color = dyad)) + 
  geom_line(aes(y = cumulativerateOther, color = dyad)) +
  geom_line(data = mean_df, aes(y = mean_cumulativerateSelf), color = "red", size = 1.5) +
  geom_line(data = mean_df, aes(y = mean_cumulativerateOther), color = "blue", size = 1.5) +
  theme_classic() +
  ggtitle("Win-shift-lose-stay agent vs Alternating agent") +
  theme(legend.position = "none")  # remove the legend for clarity
```



